### 2023-12-01 Outputs/Presentation/bell_conference_slides.qmd
Repo set up

---
title: "Torrential Twitter: Climate Change, Female Politicians, and Harassment"
author: "Inessa De Angelis"
format:
  beamer:
    aspectratio: 169
    theme: Madrid
    colortheme: DarkBlue
    institute: The University of Toronto
    # toc: true
    # Table of contents not generating at the moment
editor: visual
header-includes: 
- \definecolor{DarkBlue}{rgb}{0.05, 0.15, 0.3}
- \setbeamercolor{structure}{fg=DarkBlue}
---

# Introduction

Hello!

------------------------------------------------------------------------

# Slide 2